---
title: "Probar"
author: "FULL_CODE_NEW"
date: "17/02/2021"
output: html_document
---

#Importing data

The necessary data are the pollen percentages of each MODERN pollen sample and the biome observed in the search window on the PNV map.

Importing pollen percentages
*Chunk 2*
```{r}
#   Modify path to the data
polen_path <- "/Users/esmer/UoR/PhD/Intra_biome_variation_method_paper_code/Intra_biome_variation_method/Inputs/Modern_and_fossil_pollen_percentages/"
polen_name <- "Modern_pollen_perc_emb_smpds_June2021.csv"
polen_file <- paste(polen_path, polen_name, sep="")
polen <- read.csv(polen_file) 

str(polen)
```



Importing observed biome by using an particular search window
*Chunk 3*
```{r}
#   Modify path to the data
pnv_path <- "/Users/esmer/UoR/PhD/Intra_biome_variation_method_paper_code/Intra_biome_variation_method/Inputs/Extract_PNV_biomes/"
pnv_name <- "resolution1kmfrom250m_window21km_perside.csv"
pnv_file <- paste(pnv_path, pnv_name, sep="")
pnv_obs <- read.csv(pnv_file) 

pnv_obs <- pnv_obs %>% select(-c(Dominant_num,Subdominant_num,Third,Third_num))

str(pnv_obs)
```


#Setting some traits
*Chunk 4*
```{r}
#Set path to output files
path_to_results <- "/Users/esmer/UoR/PhD/Intra_biome_variation_method_paper_code/Intra_biome_variation_method/Outputs/Comparison_matrices_and_metrics/"

#Set prefix name of the iteration
iteration <- "EC4c_f_ep0.5"

#Set proportion of the data to be used as training dataset (0.75 for a training/testing ratio of 75:25)
SampFrac <- 0.7

#Set value for epsilon
EpsVal<- 0.5
```
#Filtering the data
  For each pollen sample, join the percentages of pollen with the observed biome in a search window on the PNV map. Keep only one sample for each pollen sample location by random sampling. Discard samples whose observed biome is *NA*. 
*Chunk 5*
```{r}
polen_wide <- polen %>% 
  pivot_wider(names_from = taxon_name, values_from = taxon_percent, values_fill = 0)

polen_wide_o<-pnv_obs %>% 
  inner_join(polen_wide, by="ID_SAMPLE") #join pollen percentages and observed biome

set.seed(2501)
ambwide_amg <- polen_wide_o %>%
  group_by(latitude,longitude) %>% 
  slice_sample(n=1) %>% #keep only one sample per location
  ungroup() %>% 
  subset(!is.na(Dominant)) #Remove samples with NA observed biome

#This is the total number of samples left in each biome
ambwide_amg %>% 
  ggplot(aes(Dominant)) +
  geom_bar()+
  geom_text(stat='count', aes(label=..count..), vjust=-0.5)
```



Downsampling any majoraty class to have the same size of the medium size class. In this case we downsampled any majority class towards GRAM by using Themis R package. https://themis.tidymodels.org/

```{r}
ratio <- ambwide_amg %>% count(Dominant) %>% 
  mutate(med = median(n)) %>% 
  pivot_wider(names_from = Dominant, values_from = n) %>% 
  select(med,GRAM,DESE) %>%
  #mutate(ratioo = med/COOL) %>% 
  mutate(ratioo = GRAM/DESE) %>%
  select(ratioo) %>% as_vector()

#Random downsampling majority classes
set.seed(2468)
  recipes::recipe(~., ambwide_amg) %>%
  themis::step_downsample(Dominant, under_ratio = ratio) %>%
  recipes::prep() %>%
  recipes::bake(new_data = NULL) -> oversamp
```


Split training and testing data
```{r}
train_data <- oversamp %>% 
  group_by(Dominant) %>% 
  sample_frac(SampFrac)

train_data %>% 
  ggplot(aes(Dominant)) +
  geom_bar()+
  geom_text(stat='count', aes(label=..count..), vjust=-0.5)

test_data <- anti_join(oversamp,train_data)

test_data %>% 
  ggplot(aes(Dominant)) +
  geom_bar()+
  geom_text(stat='count', aes(label=..count..), vjust=-0.5)
```


Get the TRAINING data table for the algorithm
```{r}
train_data <- train_data %>% 
   pivot_longer(!c(ID_SAMPLE,entity_name,latitude,longitude,
                  Dominant,Subdominant), 
               names_to = "taxon_name", values_to = "taxon_percent") %>% 
  group_by(Dominant,taxon_name) %>%
  #filter taxa that only occurs once in each group, since we cannot calculate the standard deviation
  filter(n()>=2) %>% 
  ungroup() %>%
  pivot_wider(names_from = taxon_name, values_from = taxon_percent, values_fill = 0) %>% 
  pivot_longer(!c(ID_SAMPLE,entity_name,latitude,longitude,
                  Dominant,Subdominant), 
               names_to = "taxon_name", values_to = "taxon_percent") #This is the object needed for the box-plots
```




```{r}
train_data %>% 
  subset(train_data$Dominant=="GRAM") -> aux1

train_data %>% 
  subset(train_data$Dominant=="GRAM") %>% 
  group_by(taxon_name) %>% 
  summarise_at(vars(taxon_percent),              # Specify column
               list(mean = mean)) %>%       # Specify function
  top_n(12, wt=mean) %>% 
  ungroup() %>% 
  dplyr::select(taxon_name) %>%
  distinct() %>% 
  inner_join(aux1) %>% 
  ggplot(aes(x=reorder(taxon_name, -taxon_percent), y=taxon_percent)) +
  geom_boxplot(outlier.alpha = 0.25)+
  expand_limits(y = c(0, 100))+
  scale_y_continuous(breaks=seq(0, 100, 20))+
  annotate(geom = 'text', x = 11.5, y = 95, label = "GRAM", size=4)+
  labs(x="", y="Abundance (%)", title = "a")+
  theme(axis.text.x = element_text(face = "italic",size=14, angle = 90, hjust = 1, vjust = 0.3),
        axis.text.y = element_text(size=12),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14, angle=90),
        plot.caption = element_text(size=14,hjust = 0, face = "bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA))->gram

gram

#ggsave("/My Drive/Github_esme/Within_variability_technique/Output/Biomes_boxplots/GRAM_EC4c_SHORT.tiff", plot = last_plot(), width = 6, height = 5, dpi = 600)
```



```{r}
# folder where you want the graphs to be saved:
results <- "/Users/esmer/UoR/PhD/Intra_biome_variation_method_paper_code/Intra_biome_variation_method/Outputs/Biomes_boxplots/"  


  # create list of counties in data to loop over 
  biome_list <- unique(train_data$Dominant)
  
  
  # create for loop to produce ggplot2 graphs 
  for (i in seq_along(biome_list)) { 
    
  aux1<-train_data %>% 
        subset(train_data$Dominant==biome_list[i])
      
 plot<-train_data %>% 
        subset(train_data$Dominant==biome_list[i]) %>% 
  group_by(taxon_name) %>% 
  summarise_at(vars(taxon_percent),              # Specify column
               list(mean = mean)) %>%       # Specify function
  top_n(25, wt=mean) %>% 
  ungroup() %>% 
  dplyr::select(taxon_name) %>%
  distinct() %>% 
  inner_join(aux1) %>% 
  ggplot(aes(x=reorder(taxon_name, -taxon_percent), y=taxon_percent)) +
  geom_boxplot(outlier.alpha = 0.25)+
  expand_limits(y = c(0, 100))+
  scale_y_continuous(breaks=seq(0, 100, 20))+
  labs(x="Taxon", y="Abundance (%)", title = biome_list[i])+
  theme(axis.text.x = element_text(face = "italic",size=11,angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size=11),
        axis.title.x = element_text(size=11),
        axis.title.y = element_text(size=11, angle=90),
        plot.caption = element_text(size=12,hjust = 0, face = "bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA))

    ggsave(plot, file=paste(results, 
                           biome_list[i], "_EC4c.png", sep=''),scale = 1, dpi = 600)
    
    print(plot)
  }
#}
```

