---
title: "PROB_DOWNCORE"
author: "FULL_CODE_NEW"
date: "02/12/2020"
output: html_document
---

#Importing data

The necessary data are the pollen percentages of each MODERN pollen sample and the biome observed in the search window on the PNV map.

Importing pollen percentages
*Chunk 2*
```{r}
#   Modify path to the data
polen_path <- "/Users/esmer/UoR/PhD/Intra_biome_variation_method_paper_code/Intra_biome_variation_method/Inputs/Modern_and_fossil_pollen_percentages/"
polen_name <- "Modern_pollen_perc_emb_smpds_June2021.csv"
polen_file <- paste(polen_path, polen_name, sep="")
polen <- read.csv(polen_file) 

str(polen)
```



Importing observed biome by using an particular search window
*Chunk 3*
```{r}
#   Modify path to the data
pnv_path <- "/Users/esmer/UoR/PhD/Intra_biome_variation_method_paper_code/Intra_biome_variation_method/Inputs/Extract_PNV_biomes/"
pnv_name <- "resolution1kmfrom250m_window21km_perside.csv"
pnv_file <- paste(pnv_path, pnv_name, sep="")
pnv_obs <- read.csv(pnv_file) 

pnv_obs <- pnv_obs %>% select(-c(Dominant_num,Subdominant_num,Third,Third_num))

str(pnv_obs)
```


#Setting some traits
*Chunk 4*
```{r}
#Set prefix name of the iteration
iteration <- "EC4c_f_ep0.5"

#Set proportion of the data to be used as training dataset (0.75 for a training/testing ratio of 75:25)
SampFrac <- 0.7

#Set value for epsilon
EpsVal<- 0.5
```
#Filtering the data
  For each pollen sample, join the percentages of pollen with the observed biome in a search window on the PNV map. Keep only one sample for each pollen sample location by random sampling. Discard samples whose observed biome is *NA*. 
*Chunk 5*
```{r}
polen_wide <- polen %>% 
  pivot_wider(names_from = taxon_name, values_from = taxon_percent, values_fill = 0)

polen_wide_o<-pnv_obs %>% 
  inner_join(polen_wide, by="ID_SAMPLE") #join pollen percentages and observed biome

set.seed(2501)
ambwide_amg <- polen_wide_o %>%
  group_by(latitude,longitude) %>% 
  slice_sample(n=1) %>% #keep only one sample per location
  ungroup() %>% 
  subset(!is.na(Dominant)) #Remove samples with NA observed biome
```


Downsampling any majoraty class to have the same size of the medium size class. In this case we downsampled any majority class towards GRAM
```{r}
ratio <- ambwide_amg %>% count(Dominant) %>% 
  mutate(med = median(n)) %>% 
  pivot_wider(names_from = Dominant, values_from = n) %>% 
  select(med,GRAM,DESE) %>%
  #mutate(ratioo = med/COOL) %>% 
  mutate(ratioo = GRAM/DESE) %>%
  select(ratioo) %>% as_vector()

#Random downsampling majority classes
set.seed(2468)
  recipes::recipe(~., ambwide_amg) %>%
  themis::step_downsample(Dominant, under_ratio = ratio) %>%
  recipes::prep() %>%
  recipes::bake(new_data = NULL) -> oversamp


#Split training data
train_data <- oversamp %>% 
  group_by(Dominant) %>% 
  sample_frac(SampFrac)
```


Get the TRAINING data table for the algorithm
```{r}
train_data <- train_data %>% 
   pivot_longer(!c(ID_SAMPLE,entity_name,latitude,longitude,
                  Dominant,Subdominant), 
               names_to = "taxon_name", values_to = "taxon_percent") %>% 
  group_by(Dominant,taxon_name) %>%
  #filter taxa that only occurs once in each group, since we cannot calculate the standard deviation
  filter(n()>=2) %>% 
  ungroup() %>%
  pivot_wider(names_from = taxon_name, values_from = taxon_percent, values_fill = 0) %>% 
  pivot_longer(!c(ID_SAMPLE,entity_name,latitude,longitude,
                  Dominant,Subdominant), 
               names_to = "taxon_name", values_to = "taxon_percent") #This is the object needed for the box-plots

train_data <- train_data %>% 
  ungroup() %>% 
  group_by(Dominant, taxon_name) %>%  # Specify group indicator
  summarise_at(vars(taxon_percent),list(Mean = mean,
                                        Stdev = sd)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = Dominant, values_from = c(Mean, Stdev))
```


```{r}
#Keep only the obtained train data
rm(list = setdiff(ls(),"train_data"))


DowncoreData <- read.csv("/Users/esmer/UoR/PhD/Intra_biome_variation_method_paper_code/Intra_biome_variation_method/Inputs/Modern_and_fossil_pollen_percentages/SampleData_DOWNCORE_June2021.csv",header=TRUE,sep=",")

#Keep only modern samples
DowncoreData <- DowncoreData %>% subset(age<=150)


#Create a table AllData
alldata <- full_join(DowncoreData, train_data,by="taxon_name")
str(alldata)
```


##PREDICTED BIOME
Prepare the data
```{r}
alldata <- alldata %>% 
  ##Remove taxa that are in the training set, but do not appear in the downcore data of the EMBSECBIO region
  filter(!is.na(ID_SAMPLE)) %>% 
  subset(!is.na(Mean_CENF)) #Remove taxa that are in the testing set, but not in the training set.
```


###Dissimilarity index
```{r}
sqsc <- alldata %>% 
  mutate(Epsilon=0.5)%>% #Set a value for epsilon
  mutate(TUND_Sqrt=(taxon_percent-Mean_TUND)^2/((Stdev_TUND+Epsilon)^2)) %>% 
  mutate(DESE_Sqrt=(taxon_percent-Mean_DESE)^2/((Stdev_DESE+Epsilon)^2)) %>% 
  mutate(GRAM_Sqrt=(taxon_percent-Mean_GRAM)^2/((Stdev_GRAM+Epsilon)^2)) %>% 
  mutate(XSHB_Sqrt=(taxon_percent-Mean_XSHB)^2/((Stdev_XSHB+Epsilon)^2)) %>% 
  mutate(CENF_Sqrt=(taxon_percent-Mean_CENF)^2/((Stdev_CENF+Epsilon)^2)) %>% 
  mutate(TEDE_Sqrt=(taxon_percent-Mean_TEDE)^2/((Stdev_TEDE+Epsilon)^2)) %>% 
  mutate(CMIX_Sqrt=(taxon_percent-Mean_CMIX)^2/((Stdev_CMIX+Epsilon)^2)) %>% 
  mutate(ENWD_Sqrt=(taxon_percent-Mean_ENWD)^2/((Stdev_ENWD+Epsilon)^2)) %>% 
  mutate(WTSFS_Sqrt=(taxon_percent-Mean_WTSFS)^2/((Stdev_WTSFS+Epsilon)^2)) %>% 
  dplyr::select(ID_SAMPLE,age,TUND_Sqrt,DESE_Sqrt,GRAM_Sqrt,XSHB_Sqrt,CENF_Sqrt,
         TEDE_Sqrt,CMIX_Sqrt,ENWD_Sqrt,WTSFS_Sqrt) %>% 
  group_by(ID_SAMPLE,age) %>% 
  summarise(across(c(TUND_Sqrt,DESE_Sqrt,GRAM_Sqrt,XSHB_Sqrt,CENF_Sqrt,
         TEDE_Sqrt,CMIX_Sqrt,ENWD_Sqrt,WTSFS_Sqrt), sum))%>% 
  mutate(across(c(TUND_Sqrt,DESE_Sqrt,GRAM_Sqrt,XSHB_Sqrt,CENF_Sqrt,
         TEDE_Sqrt,CMIX_Sqrt,ENWD_Sqrt,WTSFS_Sqrt), sqrt)) %>% 
  ungroup()
```


###Get recirpocal of the scores and normalization
```{r}
biomes <- sqsc %>% 
  mutate(TUND=exp(-TUND_Sqrt/100)) %>% 
  mutate(DESE=exp(-DESE_Sqrt/100)) %>% 
  mutate(GRAM=exp(-GRAM_Sqrt/100)) %>% 
  mutate(XSHB=exp(-XSHB_Sqrt/100)) %>% 
  mutate(ENWD=exp(-ENWD_Sqrt/100)) %>% 
  mutate(WTSFS=exp(-WTSFS_Sqrt/100)) %>% 
  mutate(CENF=exp(-CENF_Sqrt/100)) %>% 
  mutate(CMIX=exp(-CMIX_Sqrt/100)) %>% 
  mutate(TEDE=exp(-TEDE_Sqrt/100)) %>%   
  dplyr::select(ID_SAMPLE,age,TUND,DESE,GRAM,XSHB,ENWD,
         WTSFS,CENF,CMIX,TEDE)
```



For maps
```{r}
alldatafjoin <- alldata %>% 
  dplyr::select(entity_name,latitude,longitude,ID_SAMPLE) %>% 
  distinct()

biomes_maps <- biomes %>% 
  mutate(predicted_biome=colnames(biomes [,3:11])[apply(biomes [,3:11], 1, which.max)]) %>% 
  mutate(predicted_score=apply(biomes [,3:11], 1, max)) %>% 
  dplyr::select(ID_SAMPLE,age,predicted_biome, predicted_score) %>% 
  inner_join(alldatafjoin, by="ID_SAMPLE")
```


```{r}
colores<- c("XSHB"="#E6706E","GRAM"="#FAD272","DESE"="#F7FFCA","ENWD"="#E0E376","WTSFS"="#8FB83E","COOL"="#BBA6D0","CENF"="#6B2F85","XSHB"="#E6706E","TEDE"="#005B89","CMIX"="#28B8C8","TUND"="#B03E8D","CDFO"="#BBA6D0")
```



```{r}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
#class(world)
#land<-raster::shapefile("/Users/Esmesaurio/OneDrive - University of Reading/PAPER_NEW_BIOMISATION_FIGURES/Natural Earth Data/Medium_scale/ne_50m_land/ne_50m_land.shp")

ocean<- raster::shapefile("E:/UoR Biomisation/PAPER_NEW_BIOMISATION_FIGURES/Natural Earth Data/Medium_scale/ne_50m_ocean/ne_50m_ocean.shp")

elevation<- raster::raster("E:/UoR Biomisation/PAPER_NEW_BIOMISATION_FIGURES/Natural Earth Data/Medium_scale/MSR_50M_manual_elevation/MSR_50M/MSR_50M.tif")

#  Define your own domine:
quicksubset <- function(x, longlat){
# longlat should be a vector of four values: c(xmin, xmax, ymin, ymax)
  x@data$id <- rownames(x@data)
  xf = fortify(x, region="id")
  xjoin = plyr::join(xf, x@data, by="id")
  xsubset <- subset(xjoin, xjoin$long > longlat[1] & xjoin$long < longlat[2] &
                           xjoin$lat > longlat[3] & xjoin$lat < longlat[4])
  xsubset
}
 
#ylim = c(28.05317,49.2431), xlim = c(19.54792, 62.44992)
domain <- c(19.54792, 63, 28.05317,49.2431)
#landsubset <- quicksubset(land, domain)
 
natelevation <- raster::crop(elevation, y=raster::extent(domain))

#rast_table <- data.frame(xyFromCell(natelevation, 1:ncell(natelevation)),
                         #getValues(natelevation/255))

raster::ratify(natelevation) %>% 
  #convert the raster to points for plotting
  raster::rasterToPoints() %>% 
  #Make the points a dataframe for ggplot
  data.frame()->natelevation

#Make appropriate column headings
colnames(natelevation) <- c("Longitude", "Latitude", "elevation")
```



```{r}
#coord_sf(ylim = c(40.3,43.65), xlim = c(39.2, 46.5), expand = FALSE)+

plot_b <- ggplot(data = world) +
  #geom_sf(col="black",fill="transparent") + #Using if we want political borders
  coord_sf(ylim = c(28.05317, 49.2431), xlim = c(19.88,62), expand = FALSE) +
  #coord_sf(ylim = c(28.05317,49.2431), xlim = c(19.54792, 63), expand = FALSE)+
  geom_raster(data=natelevation, aes(y=Latitude, x=Longitude, fill=elevation), show.legend = F)+
  scale_fill_gradientn(colours = c("#6c8396", "#fcfcfc"))+ #f8f8f8 en lugar de fcfcfc  
  geom_polygon(data = ocean, aes(x=long, y=lat, group=group), fill= "white")+#E2F1FC '#E2F1FC'
  geom_point(data = biomes_maps, mapping = aes(x = longitude, y = latitude, colour=predicted_biome, size=predicted_score), alpha=1)+
      scale_size(limits = c(0.925,0.999),range = c(0.0001,2.5),breaks = c(0.94,0.96,0.98)) + #or: c(0.1,5)
      #scale_size(limits = c(min(slig$predicted_score),max(slig$predicted_score)),range = c(1,5),breaks = c(0.1,0.3))+ #or: c(0.1,5)
      scale_color_manual(values = colores)+
  geom_rect(xmin = 39.2, xmax = 46.5,   ymin = 40.3, ymax = 43.65,  color="black", fill = "transparent", linetype="dashed", size=0.15) +
  annotate(geom = "text", x = 34.25, y = 43.25, label = "Black Sea", fontface = "italic", color = "#666A6A", size = 2) +
  annotate(geom = "text", x = 28, y = 33.5, label = "Mediterranean Sea", fontface = "italic", color = "#666A6A", size = 2) +
  annotate(geom = "text", x = 50.3, y = 42.5, label = "Caspian", fontface = "italic", color = "#666A6A", size = 2) +
  annotate(geom = "text", x = 50.65, y = 41.2, label = "Sea", fontface = "italic", color = "#666A6A", size = 2) +
  labs(x = "Longitude", y = "Latitude", subtitle = "(b)") +
  guides(color = FALSE) +
  theme(plot.subtitle= element_text(size=8, face="bold", color="black"),
        axis.title.x = element_blank(),
        #axis.title.x = element_text(size=8,color = "black"),
        axis.title.y = element_blank(),
        #axis.title.y = element_text(size=8, angle=90),
        axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=8),
        panel.grid.major = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        legend.position = c(0.945, 0.095),
        legend.background = element_blank(),
        legend.key.width = unit(0.25,"cm"),
        legend.key.height = unit(0.25,"cm"),
        legend.key = element_rect(fill="transparent",color = "transparent"),
        legend.text = element_text(size=6),
        legend.title = element_blank(),
        legend.spacing.x = unit(0, 'cm'),
        legend.spacing.y = unit(0,"cm"))

plot_b

#ggsave(file="C:/Users/esmer/UoR/PhD/PAPER_RESUBMISION_3RD/Paper_figure_Modern_point_maps_plot/modern_map_points.pdf", plot = last_plot(), units = "cm", height = 6.021, width = 12, dpi = 85)
```



```{r}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
#class(world)
#land<-raster::shapefile("/Users/Esmesaurio/OneDrive - University of Reading/PAPER_NEW_BIOMISATION_FIGURES/Natural Earth Data/Medium_scale/ne_50m_land/ne_50m_land.shp")

ocean<- raster::shapefile("E:/UoR Biomisation/PAPER_NEW_BIOMISATION_FIGURES/Natural Earth Data/Medium_scale/ne_50m_ocean/ne_50m_ocean.shp")

elevation<- raster::raster("E:/UoR Biomisation/PAPER_NEW_BIOMISATION_FIGURES/Natural Earth Data/Large_scale/Elevation_GRAY_HR_SR/GRAY_HR_SR.tif")

#  Define your own domine:
quicksubset <- function(x, longlat){
# longlat should be a vector of four values: c(xmin, xmax, ymin, ymax)
  x@data$id <- rownames(x@data)
  xf = fortify(x, region="id")
  xjoin = plyr::join(xf, x@data, by="id")
  xsubset <- subset(xjoin, xjoin$long > longlat[1] & xjoin$long < longlat[2] &
                           xjoin$lat > longlat[3] & xjoin$lat < longlat[4])
  xsubset
}
 
#ylim = c(28.05317,49.2431), xlim = c(19.54792, 62.44992)
domain <- c(19.54792, 64, 28.05317,49.2431)
#landsubset <- quicksubset(land, domain)
 
natelevation <- raster::crop(elevation, y=raster::extent(domain))

#rast_table <- data.frame(xyFromCell(natelevation, 1:ncell(natelevation)),
                         #getValues(natelevation/255))

raster::ratify(natelevation) %>% 
  #convert the raster to points for plotting
  raster::rasterToPoints() %>% 
  #Make the points a dataframe for ggplot
  data.frame()->natelevation

#Make appropriate column headings
colnames(natelevation) <- c("Longitude", "Latitude", "elevation")
```



```{r}
#coord_sf(ylim = c(40.3,43.65), xlim = c(39.2, 46.5), expand = FALSE)+

plot_c <- ggplot(data = world) +
  #geom_sf(col="black",fill="transparent") + #Using if we want political borders
  coord_sf(ylim = c(40.3,43.65), xlim = c(39.2, 46.5), expand = FALSE)+
  geom_raster(data=natelevation, aes(y=Latitude, x=Longitude, fill=elevation), show.legend = F)+
  scale_x_continuous(breaks = c(40,42,44,46)) +
  scale_y_continuous(breaks = c(41,42,43)) +
  scale_fill_gradientn(colours = c("#6c8396", "#fcfcfc"))+ #f8f8f8 en lugar de fcfcfc
  geom_polygon(data = ocean, aes(x=long, y=lat, group=group), fill= "white")+#E2F1FC '#E2F1FC'
  geom_point(data = biomes_maps, mapping = aes(x = longitude, y = latitude, colour=predicted_biome, size=predicted_score), alpha=1)+
      scale_size(limits = c(0.91,0.999),range = c(0.01,4),breaks = c(0.94,0.96,0.98)) + #or: c(0.1,5)
      scale_color_manual(values = colores)+
  labs(x = "Longitude", y = "Latitude", subtitle = "(c)") +
  guides(color = FALSE) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=8),
        panel.grid.major = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        legend.position = c(0.94, 0.085),
        legend.background = element_blank(),
        legend.key.width = unit(0.25,"cm"),
        legend.key.height = unit(0.25,"cm"),
        legend.key = element_rect(fill="transparent",color = "transparent"),
        legend.text = element_text(size=6),
        legend.title = element_blank(),
        legend.spacing.x = unit(0, 'cm'),
        legend.spacing.y = unit(0,"cm"))

plot_c

#ggsave(file="C:/Users/esmer/UoR/PhD/PAPER_RESUBMISION/Modern_point_maps_plot/modern_map_closeup.pdf", plot = last_plot(), units = "cm", height = 6.021, width = 12, dpi = 85)
```


```{r}
# read potential natural vegetation data sage_veg30.nc:
# modify the following path to reflect local files
vegtype_path <- "E:/UoR Biomisation/NEW_ITERATIONS_JANUARY_2021/RASTERS/netcdf files/"
vegtype_name <- "EMBSECBIO_5km_RasterToNetCDF.nc" #the map in the paper was constructed by using the twoPointFivekm file
vegtype_file <- paste(vegtype_path, vegtype_name, sep="")
vegtype <- raster::raster(vegtype_file, varname="biome_type")

#plot(vegtype)
```


```{r}
#rm(list=ls()) #clear workspace
#Ratify values as categorical
raster::ratify(vegtype) %>% 
  raster::rasterToPoints() %>% #convert the raster to points for plotting
  data.frame()->df #Make the points a dataframe for ggplot

colnames(df) <- c("Longitude", "Latitude", "biome_type") #Make appropriate column headings

df %>% filter(Longitude<63)->df

df$biome_type[df$biome_type==8]<-9 #Amalgamating types of COOL

df$biome_type[df$biome_type==16] <- 20 #Amalgamating types of TUND
#df %>% 
  #dplyr::select(biome_type) %>% 
  #distinct() %>% 
  #arrange(biome_type)

#Call in point data, in this case a fake transect (csv file with lat and lon coordinates)
#sites <- data.frame(read.csv(“/your/path/to/pointfile.csv”))
```


```{r}
#tundra mas oscuro
cols<-  c("4"="#8FB83E", "8"="#BBA6D0", "9"="#28B8C8", "13"="#005B89", "14"="#BBA6D0", "15"="#6B2F85","16"="#E6706E","17"="#E0E376", "20"="#E6706E", "22"="#FAD272","27"="#F7FFCA", "28"="#B03E8D")#, "30"="#39c9bb","31"="#39c9bb")


nams<- c("20"="XSHB","22"="GRAM","27"="DESE","17"="ENWD","4"="WTSFS","8"="COOL","15"="CENF","16"="XSHB","13"="TEDE","9"="CMIX","28"="TUND","14"="CDFO") #"30"="TUND", "31"="TUND", "14"="CDFO")
```


```{r}
#Now make the map
plot_a <- ggplot(data = world) +
    geom_sf(col="transparent",fill="transparent", size=0.75) +
    coord_sf(ylim = c(28.05317, 49.2431), xlim = c(19.88,62), expand = FALSE) +
    #geom_tile(data=df, aes(y=Latitude, x=Longitude, fill=factor(biome_type))) +
    #scale_fill_manual(values = cols, name="Biome type", labels=nams) +
    ggspatial::annotation_scale(location = "br", width_hint = 0.080, height = unit(0.18, "cm"), pad_x = unit(0.15, "cm"), pad_y = unit(0.15, "cm")) +
    ggspatial::annotation_north_arrow(location = "tr", which_north = "true",
        height = unit(0.40, "cm"),
        width = unit(0.30, "cm"),
        pad_x = unit(0.20, "cm"), pad_y = unit(0.20, "cm")) +
        #style = north_arrow_orienteering) +
  annotate(geom = "text", x = 34.25, y = 43.25, label = "Black Sea", fontface = "italic", color = "#666A6A", size = 2) +
  annotate(geom = "text", x = 26, y = 34, label = "Mediterranean Sea", fontface = "italic", color = "#666A6A", size = 2) +
  annotate(geom = "text", x = 50.2, y = 42.5, label = "Caspian", fontface = "italic", color = "#666A6A", size = 2) +
  annotate(geom = "text", x = 50.65, y = 41.3, label = "Sea", fontface = "italic", color = "#666A6A", size = 2) +
  guides(fill = FALSE) + 
  labs(x = "Longitude", y = "Latitude", subtitle = "(a)") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        #axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=8),
        panel.grid.major = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        legend.position = c(0.94, 0.085),
        legend.background = element_blank(),
        legend.key.width = unit(0.25,"cm"),
        legend.key.height = unit(0.25,"cm"),
        legend.key = element_rect(fill="transparent",color = "transparent"),
        legend.text = element_text(size=6),
        legend.title = element_blank(),
        legend.spacing.x = unit(0, 'cm'),
        legend.spacing.y = unit(0,"cm"))
  
  
plot_a

#ggsave(file="C:/Users/esmer/UoR/PhD/PAPER_RESUBMISION/Modern_point_maps_plot/modern_map_raster_outside.pdf", plot = last_plot(), units = "cm", height = 6.021, width = 12, dpi = 85)
```



